\documentclass[11pt]{article}
\usepackage[margin=1in, top=15pt]{geometry}
\usepackage{amsmath, amssymb, xcolor, booktabs, graphicx, tikz, enumitem}
\usetikzlibrary{graphs, graphs.standard, quotes}
\title{CS 225}
\author{Graph Theory}
\date{Fall 2020}

\begin{document}
\maketitle
\hrule

% \begin{center}\underline{\bf \large Definitions}\end{center}
\bigskip

{\bf Graph:} A graph $G$ is a nonempty set $V(G)$ of vertices,
and a set $E(G)$ of edges, where each edge is associated with a 
set of either one or two vertices, called the {\bf endpoints} 
of that edge. 

\bigskip
Here is a graph vith $V(G) = \{v_1, v_2\}$ and $E(G) = \{e_1\} = \{\{v_1, v_2\}\}$: 

\begin{center}
\begin{tikzpicture}
\graph {
    v1/$v_1$ -- ["$e_1$"] v2/$v_2$;
};
\end{tikzpicture}
\end{center}

\bigskip 
The {\bf edge to endpoint} function sends an edge to its set of endpoints. 
If we define an edge as the set of its endpoints (which we can do for simple 
graphs), this is pretty straightforward. 

\bigskip
An edge with just one endpoint is a loop (note that the loop drawn here is 
directed due to how my drawing program is working at the moment. Please 
ignore the arrow on the loop edge):
\begin{center}
    \begin{tikzpicture}
        \graph {
        v1/$v_1$ -- ["$e_1$", loop, swap] v1;
        };
    \end{tikzpicture}
\end{center}

Two or more {\it distinct} edges with the same set of endpoints are said 
to be {\bf parallel}:

\begin{center}
    \begin{tikzpicture}
        \graph {
        v1/$v_1$ -- ["$e_1$", bend left] v2/$v_2$;
        v1/$v_1$ -- ["$e_2$", bend right, swap] v2/$v_2$;
        };
    \end{tikzpicture}
\end{center}

Two vertices connected by an edge are adjacent. In the picture below, $v_1$ 
and $v_2$ are adjacent vertices, and $v_2$ is adjacent to itself.

\begin{center}
\begin{tikzpicture}
\graph {
    v1/$v_1$ -- ["$e_1$"] v2/$v_2$;
    v2 -- [loop right, "$e_2$"] v2;
};
\end{tikzpicture}
\end{center}

An edge is {\bf incident} on each of its endpoints, and two edges incident 
to the same endpoint are {\bf adjacent}. Below, edges $e_1$ and $e_2$ are 
adjacent since they share the endpoint $v_2$. 
\begin{center}
    \begin{tikzpicture}
        \graph {
        v1/$v_1$ -- ["$e_1$"] v2/$v_2$ -- ["$e_2$"] v3/$v_3$;
        };
    \end{tikzpicture}
\end{center}

An isolated vertex:
\begin{center}
    \begin{tikzpicture}
        \graph {
          v1/$v_1$ -- v1/$v_1$;  
        };
    \end{tikzpicture}
\end{center}

\newpage
\newgeometry{top=1in}
A {\bf directed graph} has vertices and edges, but the edges are now ordered 
pairs. We can represent a directed graph as vertices and edges where the edges 
have arrows. Below is a graph with vertices $V(G) = \{v_1, v_2\}$ and directed 
edges $D(G) = \{e_1\} = \{(v_1, v_2)\}$. 

\begin{center}
    \begin{tikzpicture}
        \graph {
         v1/$v_1$ -> ["$e_1$"] v2/$v_2$;   
        };
    \end{tikzpicture}
\end{center}

If $G$ is a graph and $v$ is a vertex of $G$, then the {\bf degree} of $v$, 
$\deg(v)$ is the number of edges that are incident on $v$, with an edge that 
is a loop counted twice. Below $v_1$ has degree 1 and $v_2$ has degree 3:
\begin{center}
    \begin{tikzpicture}
        \graph {
         v1/$v_1$ -- ["$e_1$"] v2/$v_2$ -- [loop right] v2;   
        };
    \end{tikzpicture}
\end{center}

The {\bf total degree} of a graph is the sum of degrees of all vertices of the 
graph. In the graph below, the total degree is 4. 
\begin{center}
    \begin{tikzpicture}
        \graph {
         v1/$v_1$ -- ["$e_1$"] v2/$v_2$ -- [loop right] v2;   
        };
    \end{tikzpicture}
\end{center}

\bigskip
{\bf Theorem 4.9.1: The handshake theorem} If $G$ is any graph then the sum of 
degrees of all the vertices of $G$ is equal to twice the number of edges of $G$. 

\bigskip
{\bf Corollary 4.9.2:} The total degree of a graph is even. 

\bigskip
{\bf Proposition 4.9.3:} In any graph there is an even number of vertices of 
odd degree. 

(if not then Corollary 4.9.2 would be false)

\bigskip
A {\bf simple} graph is a graph that does not have any loops or parallel edges. 

\bigskip
If $n$ is a positive integer, then a {\bf complete graph on $n$ vertices}, 
denoted $K_n$, is a simple graph with $n$ vertices and exactly one edge 
connecting each pair of distinct vertices. Below are $K_5$ and $K_{10}$:

\begin{center}
    \begin{tikzpicture}
        \graph[clique, n=5, clockwise, radius=2cm, math nodes] {
            \foreach \i in {1, ..., 5}{
                v\i/v_\i,
            }
        };
    \end{tikzpicture}
\end{center}


\begin{center}
    \begin{tikzpicture}
        \graph[clique, n=10, clockwise, radius=2cm, math nodes] {
            \foreach \i in {1, ..., 10}{
                v\i/v_{\i},
            }
        };
    \end{tikzpicture}
\end{center}

{\bf Example 4.9.9} $K_n$ has $\dfrac{n(n-1)}{2}$ edges. 

\bigskip
Let $m$ and $n$ be positive integers. A {\bf complete bipartite graph on $(m,n)$ 
vertices}, denoted $K_{m, n}$, is a simple graph whose vertices are divided 
into two distinct subsets, $V$ with $m$ vertices and $W$ with $n$ vertices, 
in such a way that 
\begin{enumerate}
    \item Every vertex of $K_{m,n}$ velongs to one of $V$ or $W$ but not both
    \item There is exactly one edge from each vertex of $V$ to each vertex of $W$
    \item There is no edge from any one vertex of $V$ to any other vertex of $V$. 
    \item There is no edge from any one vertex of $W$ to any other vertex of $W$. 
\end{enumerate}

$K_{3, 2}$ is shown below:

\begin{center}
    \begin{tikzpicture}
        \graph[ math nodes] {
        {\foreach \i in {1, ..., 3}{
            v\i / v_{\i},
        }} -- [complete bipartite]
        {\foreach \i in {1,2}{
            w\i / w_{\i},
        }}
        };
    \end{tikzpicture}
\end{center}

\begin{center}
    \underline{\bf From Chapter 10:}
\end{center}

Let $G$ be a graph, and let $v$ and $w$ be vertices of $G$. A {\bf walk from 
$v$ to $w$} is a finite alternating sequence of adjacent vertices and edges of 
$G$. A walk has the form $v_0 e_1 v_1 e_2 \dots e_n v_n$. The {\bf trivial walk} 
from $v$ to $v$ consists of the single vertex $v$. 

In the graph below, $v_1 e_1 v_2 e_2 v_3$ is a walk from $v_1$ to $v_3$. Another 
walk from $v_1$ to $v_3$ is $v_1 e_3 v_3$. 

\begin{center}
    \begin{tikzpicture}
        \graph[math nodes] {
        v_1 -- ["$e_1$"] v_2 -- ["$e_2$"]  v_3;
        v_1 -- ["$e_3$", bend right, swap]  v_3;
        };
    \end{tikzpicture}
\end{center}

A {\bf trail} from $v$ to $w$ is a walk from $v$ to $w$ with no repeated edge. 
In the graph below, $v e_1 v$ is a trail from $v$ to $v$. 
\begin{center}
    \begin{tikzpicture}
        \graph[math nodes] {
        v -- ["$e_1$", loop] v;
        };
    \end{tikzpicture}
\end{center}

\bigskip
A {\bf path} from $v$ to $w$ is a trail that does not contain a repeated vertex. 
In the graph above, the trail from $v$ to $v$ $v e_1 v$ is not a path, but the 
trivial walk is a path. 

\bigskip 
A {\bf closed walk} is a walk that starts and ends at the same vertex. 

\bigskip 
A {\bf circuit} is a closed walk that contains at least one edge and does not 
contain a repeated edge. 

\bigskip
A {\bf simple circuit} is a circuit that does not have any other repeated vertex 
except the first and last. 

\bigskip
A graph $H$ is said to be a {\bf subgraph} of a graph $G$ if and only if every 
vertex in $H$ is also a vertex in $G$ (i.e. $V(H) \subseteq V(G)$), every 
edge in $H$ is also an edge in $G$, and every edge in $H$ has the same endpoints 
as it has in $G$ (i.e. $E(H) \subseteq E(G)$). 

\bigskip
Let $G$ be a graph. Two vertices $v$ and $w$ of $G$ are {\bf connected} if and 
only if there is a walk from $v$ to $w$. The graph $G$ is {\bf connected } if and only if 
all pairs of vertices in $G$ are connected. 

\bigskip
{\bf Lemma 10.1.1} Let $G$ be a graph. Then 
\begin{enumerate}[label=\alph*.]
    \item If $G$ is connected, then any two distinct vertices of $G$ can be 
    connected by a path. 
    \item If vertices $v$ and $w$ are part of a circuit in $G$ and one edge is 
    removed from the circuit, then there still exists a trail from $v$ to $w$ 
    in $G$. 
    \item If $G$ is connected and $G$ contains a circuit, then an edge of the 
    circuit can be removed without disconnecting $G$. 
\end{enumerate}

\bigskip
A {\bf connected component} of a graph is a connected subgraph of largest possible 
size. 

\bigskip 
A graph $H$ is a {\bf connected component} of a graph $G$ if and only if 
\begin{enumerate}
    \item $H$ is a subgraph of $G$.
    \item $H$ is connected. 
    \item no connected subgraph of $G$ has $H$ as a subgraph and contains 
    vertices or edges that are not in $H$. 
\end{enumerate}

\bigskip
Let $G$ be a graph. An {\bf Euler circuit} for $G$ is a circuit that contains 
every vertex and every edge of $G$. That is, an Euler circuit for $G$ is a sequence 
of adjacent vertices and edges in $G$ that has at least ne edge, starts and ends 
at the same vertex, uses every vertex of $G$ at least once, and uses every edge of 
$G$ exactly once. 

\bigskip
{\bf Theorem 10.1.2:} If a graph has an Euler circuit, then every vertex of the 
graph has positive even degree. Therefore if some vertex of a graph has odd 
degree, then the graph does not have an Euler circuit. 

\bigskip
{\bf Theorem 10.1.3:} If a graph $G$ is connected and the degree of every 
vertex of $G$ is  positive even integer, then $G$ has an Euler circuit. 

\bigskip
{\bf Theorem 10.1.4:} A graph $G$ has an Euler circuit if and only if $G$ 
is connected and every vertex of $G$ has positive even degree. 

\bigskip
Let $G$ be a graph, and let $v$ and $w$ be two distinct vertices of $G$. An 
{\bf Euler trail} from $v$ to $w$ is a sequence of adjacent edges and vertices 
that starts at $v$, ends at $w$, passes through every vertex of $G$ at least 
once, and traverses every edge of $G$ exactly once. 

\bigskip
{\bf Corollary 10.1.5:}{
    Let $G$ be a graph, and let $v$ and $w$ be two distinct vertices of $G$. 
    There is an Euler trail from $v$ to $w$ if and only if $G$ is connected, 
    $v$ and $w$ have odd degree, and all other vertices of $G$ have positive 
    even degree. 
}

\bigskip
Given a graph $G$, a {\bf Hamiltonian circuit} for $G$ is a simple circuit 
that includes every vertex of $G$. That is, a Hamiltonian circuit for $G$ is a 
sequence of adjacent vertices  and distinct edges in which every vertex of $G$ 
appears exactly once, except for the first and last, which are the same. 

\bigskip 
{\bf Proposition 10.1.6:} 
If a graph $G$ has a Hamiltonian circuit, then $G$ has a subgraph $H$ with 
the following properties:
\begin{enumerate}
    \item $H$ contains every vertex of $G$
    \item $H$ is connected 
    \item $H$ has the same number of edges as vertices. 
    \item Every vertex of $H$ has degree 2. 
\end{enumerate}

\end{document}